|

### 2021-08-09 Real Dataset SV.Rmd
Aggiornamento tabelle e commenti

SC<-matrix(nrow=2,ncol=3)
rownames(SC)<-c("Mean","Standard Deviation")
colnames(SC) <- c("Dow Jones", "S&P500", "STOXX50E")

@@

```{r include=FALSE}
Errorcomp3<-matrix(NA,ncol=2,nrow=2)
colnames(Errorcomp3)<-c("RMSE", "MAE")
rownames(Errorcomp3)<-c("N=50000", "N=100000")
Errorcomp3[1,]<-c(0.04407704,0.01667898)
Errorcomp3[2,]<-c(0.03575915,0.01476220)
```

@@

Overall, we should not overplay tiny differences.\footnote{Indeed, the mere extent of such difference also depends on the real dataset which is used for the comparison, implying different target distributions. For example, running the same experiment using the STOXX50E dataset, we have even smaller differences between the filters at N=50000, with RMS difference between the Bootstrap and the Auxiliary equal to 0.01236279 and mean absolute difference equal to 0.008028818.} Rather, we prefer to stress the points that emerged from both benchmark analyses and we can therefore support quite confidently.\
First, if on the one hand it is clear that some degree of resampling is essential, on the other we found no gains from resampling at each period.\
Second, when the number of particles is set at 5000 or more, neither benchmark could single out an algorithm that clearly outperformed the others, when restricting the analysis to procedures that feature adaptive resampling rules.\
Third, when the number of particles was set to 10, 100 or 1000, the Bootstrap particle filter and the Guided particle filter with optimal proposal kernel significantly emerged as the best performing algorithms for our SV model, suggesting that their computational cost is lower than that of the Auxiliary particle filter, which needed more particles to "get rolling".\
Finally, some discrepancy between different approximations persisted even when the number of particles was set to 50000 or 100000. However, the narrow extent of the RMSEs and MAEs with respect to both benchmarks supports the idea that at N=10000 the best-performing algorithms have already entered the phase of convergence. Even in light of these markedly more computationally expensive benchmarks, we then choose N=10000 as the number of particles which we routinely employ in this empirical application.

@@

```{r include=FALSE}
Errorcomp3<-matrix(NA,ncol=7,nrow=6)
colnames(Errorcomp3)<-c("N","BPF", "GPFOPT", "APF", "BAPF","LW", "SIS")
rownames(Errorcomp3)<-c("RMSE SP500","MAE SP500","RMSE STOXX","MAE STOXX","RMSE DJI","MAE DJI")
Errorcomp3[,1]<-c(10000,10000,10000,10000,10000,10000)
Errorcomp3[,2]<-c(0.508642137861919,0.353767519133916,0.463584874517798,0.301898325469172,0.507624385776376,0.337543646258033)
Errorcomp3[,3]<-c(0.509256032254369,0.35314674168209,0.463446053320389,0.304178601158929,0.506222076244991,0.338268970683459)
Errorcomp3[,4]<-c(0.509943715852212,0.351738603428736,0.475708272485282,0.30422270387141,0.509343139732461,0.337449244130176)
Errorcomp3[,5]<-c(0.510334461444389,0.35305864294958,0.46783471248187,0.304213895457104,0.505360139532104,0.3361502165818)
Errorcomp3[,6]<-c(0.428716693039206,0.291557534456417,0.344065170120452,0.237159281665533,0.440832171188778,0.280952286978564)
Errorcomp3[,7]<-c(0.602883459334495,0.431972177956328,0.652533564914959,0.440767074642285,0.604672122750974,0.415631855223678)

@@

Finally, we can see from Table XX the RMSE and MAE obtained with a number of particles equal to 50000, for the SP500 and the EuroStoxx50 indices. We consider the Bootstrap particle filter and the Auxiliary particle filter, which we employed in section XX; in addition to these, we add the Liu and West filter, which did not feature in that section of this work.\
In the case of the SP500 index, although the number of particles increases, the RMSE and MAE increase (although by very small amounts) for both the particle filters based on our usual specification of the SV model. Instead, for the Liu and West filter we see some improvements, although the small extent suggests that also the LW filter was already in the convergence phase with N=10000.\
In the case of the EuroStoxx50 index, while the Bootstrap filter increases both RMSE and MAE when N is raised to 50000, the APF shows improvements in its performance, according to both criteria, which instead disagree for the Liu and West filter.

```{r include=FALSE}
Errorcomp4<-matrix(NA,ncol=4,nrow=4)
colnames(Errorcomp4)<-c("N","BPF", "APF","LWF")
rownames(Errorcomp4)<-c("RMSE SP500","MAE SP500","RMSE STOXX","MAE STOXX")
Errorcomp4[,1]<-c(50000,50000,50000,50000)
Errorcomp4[,2]<-c(0.511700619467058,0.354186384429001,0.4650284,0.3034359)
Errorcomp4[,3]<-c(0.514888164962505,0.355336840861771,0.4670426,0.3040741)
Errorcomp4[,4]<-c(0.4242305,0.2906932,0.3422955,0.2395650)

>>>>

SC<-matrix(nrow=2,ncol=3)
rownames(SC)<-c("Mean","Standard Deviation")
colnames(SC) <- c("S&P500","Dow Jones","STOXX50E")

@@

```{r include=FALSE}
Errorcomp3<-matrix(NA,ncol=2,nrow=2)
colnames(Errorcomp3)<-c("RMSE", "MAE")
rownames(Errorcomp3)<-c("N=50000", "N=100000")
Errorcomp3[1,]<-c(0.015181843,0.008679359)
Errorcomp3[2,]<-c(0.013735482,0.006467285) #updated
```

@@

Overall, we should not overplay tiny differences.\footnote{Indeed, the mere extent of such difference also depends on the real dataset which is used for the comparison, implying different target distributions. For example, running the same experiment using the STOXX50E dataset, we have similar differences between the filters at N=50000, with RMS difference between the Bootstrap and the Auxiliary equal to 0.016121989 and mean absolute difference equal to 0.009306925. The extent of the difference clearly depends on the chosen parameters too: for example, at N=50000, choosing $\alpha=-0.0031$, $\beta=0.9951$ and $\tau^2=0.0074$, using SP500 data yields RMS difference equal to 0.04408 and MA difference equal to 0.01668, while using STOXX50E data we have 0.01236279 RMS difference and 0.008028818 mean absolute difference.} Rather, we prefer to stress the points that emerged from both benchmark analyses and we can therefore support quite confidently.\
First, if on the one hand it is clear that some degree of resampling is essential, on the other we found no gains from resampling at each period.\
Second, when the number of particles is set at 5000 or more, neither benchmark could single out an algorithm that clearly outperformed the others, when restricting the analysis to procedures that feature adaptive resampling rules.\
Third, when the number of particles was set to 10, 100 or 1000, the Bootstrap particle filter and the Guided particle filter with optimal proposal kernel significantly emerged as the best performing algorithms for our SV model, suggesting that their computational cost is lower than that of the Auxiliary particle filter, which needed more particles to "get rolling".\
Finally, some discrepancy between different approximations persisted even when the number of particles was set to 50000 or 100000. However, the narrow extent of the RMSEs and MAEs with respect to both benchmarks supports the idea that at N=10000 the best-performing algorithms have already entered the phase of convergence. Even in light of these markedly more computationally expensive benchmarks, we then choose N=10000 as the number of particles which we routinely employ in this empirical application.

@@

```{r include=FALSE}
# RMSE MAE comparison

library(openxlsx)
Errorcomp3<-matrix(NA,ncol=7,nrow=6)
colnames(Errorcomp3)<-c("N","BPF", "GPFOPT", "APF", "BAPF","LW", "SIS")
rownames(Errorcomp3)<-c("RMSE SP500","MAE SP500","RMSE STOXX","MAE STOXX","RMSE DJI","MAE DJI")
Errorcomp3[,1]<-c(10000,10000,10000,10000,10000,10000)

volatilitySPX<-read.xlsx("Dataset 17-21.xlsx", sheet = 1)
x1<-volatilitySPX[,3]
realisedx1<-x1
sv_bpf1<-(read.xlsx("SP500_filters_10000p.xlsx", sheet = 1))[,1]
sv_gpfopt1<-(read.xlsx("SP500_filters_10000p.xlsx", sheet = 2))[,1]
sv_apf1<-(read.xlsx("SP500_filters_10000p.xlsx", sheet = 3))[,1]
sv_lw1<-(read.xlsx("SP500_filters_10000p.xlsx", sheet = 4))[,1]
sv_bapf1<-(read.xlsx("SP500_filters_10000p.xlsx", sheet = 5))[,1]
sv_sis1<-(read.xlsx("SP500_filters_10000p.xlsx", sheet = 6))[,1]
RMSE<-function(x,xhat){sqrt(mean((x-xhat)^2))}
MAE<-function(x,xhat){mean(abs(x-xhat))}
comparablevol<-function(fun){exp(fun/2)}
Errorcomp3[1,2]<-RMSE(realisedx1,comparablevol(sv_bpf1))
Errorcomp3[1,3]<-RMSE(realisedx1,comparablevol(sv_gpfopt1))
Errorcomp3[1,4]<-RMSE(realisedx1,comparablevol(sv_apf1))
Errorcomp3[1,5]<-RMSE(realisedx1,comparablevol(sv_bapf1))
Errorcomp3[1,6]<-RMSE(realisedx1,comparablevol(sv_lw1))
Errorcomp3[1,7]<-RMSE(realisedx1,comparablevol(sv_sis1))
Errorcomp3[2,2]<-MAE(realisedx1,comparablevol(sv_bpf1))
Errorcomp3[2,3]<-MAE(realisedx1,comparablevol(sv_gpfopt1))
Errorcomp3[2,4]<-MAE(realisedx1,comparablevol(sv_apf1))
Errorcomp3[2,5]<-MAE(realisedx1,comparablevol(sv_bapf1))
Errorcomp3[2,6]<-MAE(realisedx1,comparablevol(sv_lw1))
Errorcomp3[2,7]<-MAE(realisedx1,comparablevol(sv_sis1))

volatilitySTOXX<-read.xlsx("Dataset 17-21.xlsx", sheet = 2)
x2<-volatilitySTOXX[,3]
realisedx2<-x2
sv_bpf2<-(read.xlsx("STOXX_filters_10000p.xlsx", sheet = 1))[,1]
sv_gpfopt2<-(read.xlsx("STOXX_filters_10000p.xlsx", sheet = 2))[,1]
sv_apf2<-(read.xlsx("STOXX_filters_10000p.xlsx", sheet = 3))[,1]
sv_lw2<-(read.xlsx("STOXX_filters_10000p.xlsx", sheet = 4))[,1]
sv_bapf2<-(read.xlsx("STOXX_filters_10000p.xlsx", sheet = 5))[,1]
sv_sis2<-(read.xlsx("STOXX_filters_10000p.xlsx", sheet = 6))[,1]
RMSE<-function(x,xhat){sqrt(mean((x-xhat)^2))}
MAE<-function(x,xhat){mean(abs(x-xhat))}
comparablevol<-function(fun){exp(fun/2)}
Errorcomp3[3,2]<-RMSE(realisedx2,comparablevol(sv_bpf2))
Errorcomp3[3,3]<-RMSE(realisedx2,comparablevol(sv_gpfopt2))
Errorcomp3[3,4]<-RMSE(realisedx2,comparablevol(sv_apf2))
Errorcomp3[3,5]<-RMSE(realisedx2,comparablevol(sv_bapf2))
Errorcomp3[3,6]<-RMSE(realisedx2,comparablevol(sv_lw2))
Errorcomp3[3,7]<-RMSE(realisedx2,comparablevol(sv_sis2))
Errorcomp3[4,2]<-MAE(realisedx2,comparablevol(sv_bpf2))
Errorcomp3[4,3]<-MAE(realisedx2,comparablevol(sv_gpfopt2))
Errorcomp3[4,4]<-MAE(realisedx2,comparablevol(sv_apf2))
Errorcomp3[4,5]<-MAE(realisedx2,comparablevol(sv_bapf2))
Errorcomp3[4,6]<-MAE(realisedx2,comparablevol(sv_lw2))
Errorcomp3[4,7]<-MAE(realisedx2,comparablevol(sv_sis2))

volatilityDJI<-read.xlsx("Dataset 17-21.xlsx", sheet = 3)
x3<-volatilityDJI[,3]
realisedx3<-x3
sv_bpf3<-(read.xlsx("DJI_filters_10000p.xlsx", sheet = 1))[,1]
sv_gpfopt3<-(read.xlsx("DJI_filters_10000p.xlsx", sheet = 2))[,1]
sv_apf3<-(read.xlsx("DJI_filters_10000p.xlsx", sheet = 3))[,1]
sv_lw3<-(read.xlsx("DJI_filters_10000p.xlsx", sheet = 4))[,1]
sv_bapf3<-(read.xlsx("DJI_filters_10000p.xlsx", sheet = 5))[,1]
sv_sis3<-(read.xlsx("DJI_filters_10000p.xlsx", sheet = 6))[,1]
RMSE<-function(x,xhat){sqrt(mean((x-xhat)^2))}
MAE<-function(x,xhat){mean(abs(x-xhat))}
comparablevol<-function(fun){exp(fun/2)}
Errorcomp3[5,2]<-RMSE(realisedx3,comparablevol(sv_bpf3))
Errorcomp3[5,3]<-RMSE(realisedx3,comparablevol(sv_gpfopt3))
Errorcomp3[5,4]<-RMSE(realisedx3,comparablevol(sv_apf3))
Errorcomp3[5,5]<-RMSE(realisedx3,comparablevol(sv_bapf3))
Errorcomp3[5,6]<-RMSE(realisedx3,comparablevol(sv_lw3))
Errorcomp3[5,7]<-RMSE(realisedx3,comparablevol(sv_sis3))
Errorcomp3[6,2]<-MAE(realisedx3,comparablevol(sv_bpf3))
Errorcomp3[6,3]<-MAE(realisedx3,comparablevol(sv_gpfopt3))
Errorcomp3[6,4]<-MAE(realisedx3,comparablevol(sv_apf3))
Errorcomp3[6,5]<-MAE(realisedx3,comparablevol(sv_bapf3))
Errorcomp3[6,6]<-MAE(realisedx3,comparablevol(sv_lw3))
Errorcomp3[6,7]<-MAE(realisedx3,comparablevol(sv_sis3))

@@

Finally, we can see from Table XX the RMSE and MAE obtained with a number of particles equal to 50000, for the SP500 and the EuroStoxx50 indices. We consider the Bootstrap particle filter and the Auxiliary particle filter, which we employed in section XX; in addition to these, we add the Liu and West filter, which did not feature in that section of this work.\
In the case of the SP500 index, although the number of particles increases, the RMSE and MAE increase (although by very small amounts) for both the particle filters based on our usual specification of the SV model. Instead, for the Liu and West filter we see some improvements, although the small extent suggests that also the LW filter was already in the convergence phase with N=10000.\
In the case of the EuroStoxx50 index, while the Bootstrap filter increases both RMSE and MAE when N is raised to 50000, the APF shows improvements in its performance, according to both criteria, which instead disagree for the Liu and West filter. COMMENTI DA AGGIORNARE

```{r include=FALSE}

library(openxlsx)
Errorcomp4<-matrix(NA,ncol=4,nrow=4)
colnames(Errorcomp4)<-c("N","BPF", "APF","LWF")
rownames(Errorcomp4)<-c("RMSE SP500","MAE SP500","RMSE STOXX","MAE STOXX")
Errorcomp4[,1]<-c(50000,50000,50000,50000)

volatilitySPX<-read.xlsx("Dataset 17-21.xlsx", sheet = 1)
x1<-volatilitySPX[,3]
realisedx1<-x1
sv_bpf1_5<-(read.xlsx("svbpf50000p.xlsx", sheet = 1))[,1]
sv_apf1_5<-(read.xlsx("svapf50000p.xlsx", sheet = 1))[,1]
sv_lw1_5<-(read.xlsx("svlw50000p.xlsx", sheet = 1))[,1]
RMSE<-function(x,xhat){sqrt(mean((x-xhat)^2))}
MAE<-function(x,xhat){mean(abs(x-xhat))}
comparablevol<-function(fun){exp(fun/2)}
Errorcomp4[1,2]<-RMSE(realisedx1,comparablevol(sv_bpf1_5))
Errorcomp4[1,3]<-RMSE(realisedx1,comparablevol(sv_apf1_5))
Errorcomp4[1,4]<-RMSE(realisedx1,comparablevol(sv_lw1_5))
Errorcomp4[2,2]<-MAE(realisedx1,comparablevol(sv_bpf1_5))
Errorcomp4[2,3]<-MAE(realisedx1,comparablevol(sv_apf1_5))
Errorcomp4[2,4]<-MAE(realisedx1,comparablevol(sv_lw1_5))

volatilitySPX<-read.xlsx("Dataset 17-21.xlsx", sheet = 2)
x2<-volatilitySPX[,3]
realisedx2<-x2
sv_bpf2_5<-(read.xlsx("svbpf50000p.xlsx", sheet = 2))[,1]
sv_apf2_5<-(read.xlsx("svapf50000p.xlsx", sheet = 2))[,1]
sv_lw2_5<-(read.xlsx("svlw50000p.xlsx", sheet = 2))[,1]
RMSE<-function(x,xhat){sqrt(mean((x-xhat)^2))}
MAE<-function(x,xhat){mean(abs(x-xhat))}
comparablevol<-function(fun){exp(fun/2)}
Errorcomp4[3,2]<-RMSE(realisedx2,comparablevol(sv_bpf2_5))
Errorcomp4[3,3]<-RMSE(realisedx2,comparablevol(sv_apf2_5))
Errorcomp4[3,4]<-RMSE(realisedx2,comparablevol(sv_lw2_5))
Errorcomp4[4,2]<-MAE(realisedx2,comparablevol(sv_bpf2_5))
Errorcomp4[4,3]<-MAE(realisedx2,comparablevol(sv_apf2_5))
Errorcomp4[4,4]<-MAE(realisedx2,comparablevol(sv_lw2_5))

### 2021-08-09 prova_knit_finale.Rmd
Aggiornamento tabelle e commenti

We provide both a graphical and a numerical comparison.\footnote{The sequences of filtered states shown in the graphs are obtained by using filters with 10000 particles.} From the plots it is hard to say which particle filter provides the best performances. By and large, BPF, GPF and APF behave similarly, while the Liu and West filter presents larger credible intervals but at the same time a higher coverage of the true states. On the other hand, by a RMSE comparison, the Auxiliary Particle filter provides the best performances followed bt the Liu and West filter. Moreover, notice that when the number of particles increases the accuracy improves, however not significant improvements are detected for $N>10000$.

>>>>

We provide both a graphical and a numerical comparison.\footnote{The sequences of filtered states shown in the graphs are obtained by using filters with 10000 particles.} From the plots it is hard to say which particle filter provides the best performances. By and large, BPF, GPF and APF behave similarly, while the Liu and West filter presents larger credible intervals but at the same time a higher coverage of the true states.\
On the other hand, by a RMSE comparison, we see several differences between the filters. Note that the Guided Particle Filter yields the lowest RMSE at N=100, and does not improve much even when raising the number of particles to 100000. Instead, the APF achieves the best performance in the field when N rises to 10000 and 100000, despite lagging well behind the BPF and the GPF for the three initial N thresholds. For the Bootstrap PF, the most significant improvements is that from N=100 to N=1000. We could then point at the APF as the best performing and more computationally expensive filter, while the GPF as the best one at low thresholds (\textit{i.e.} least computationally expensive, in a way). However, fine margins suggest caution in such comparative assessments.\
Rather, we would stress the performance of the Liu and West filter: indeed, it is apparent that, at least for the first four thresholds, such filter delivers the worst performance. Even at N=100000, the APF, the GPF and the BPF are much closer to each other than the LWF, which still yields the highest RMSE. Such results, in fact, do not come as a surprise: indeed, note that, while the three best performing filters are built on a model that perfectly coincides with the true DGP, the Liu and West filter considers a somewhat misspecified model, which features additional uncertainty on several parameters. Intuitively, even though the LWF features some degree of parameter learning, and would then converge to the correct figures of such parameters, the other filters "have already learnt" the true parameters from the beginning.\footnote{Such difference can be represented as follows. When trying to pin down real valued parameters, the probability that a simple initial guess for each value is correct can be deemed as null. So, in contexts where parameters are unknown, we expect the additional uncertainty featured in the LWF to prove useful, as it can adjust the initial parameters to get closer to the true unknown values. However, in synthetic contexts where we know the DGP, such uncertainty is almost useless.}\
Finally, notice that in general when the number of particles increases the accuracy improves; however, significant improvements are not detected for $N>10000$, suggesting that such threshold may be a telling one as regards the convergence of our filters.
